Cell Decompositions for Rank Two Quiver Grassmannians

### 2017-11-19 Grassmannian_cells.tex
GL_n actions

\newcommand{\bfe}{\mathbf{e}}
\newcommand{\bff}{\mathbf{f}}
\newcommand{\bfg}{\mathbf{g}}
\newcommand{\bfi}{\mathbf{i}}
\newcommand{\bfh}{\mathbf{h}}
\newcommand{\tbfe}{{\tilde\bfe}}
\newcommand{\tbff}{{\tilde\bff}}
\newcommand{\tbfg}{{\tilde\bfg}}
\newcommand{\tbfh}{{\tilde\bfh}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\cC}{\mathcal{C}}
\newcommand{\cG}{\mathcal{G}}
\newcommand{\cP}{\mathcal{P}}
\newcommand{\ui}{\underline i}
\newcommand{\uj}{\underline j}
\newcommand\udim{{\underline{\dim}\, }}
\newcommand{\pt}{\mathrm{pt}}
\newcommand{\Rep}{\mathrm{Rep}}
\newcommand{\rep}{\operatorname{rep}}
\newcommand{\Gr}{\mathrm{Gr}}
\renewcommand{\AA}{\mathbb{A}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\kk}{\Bbbk}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\Ext}{\operatorname{Ext}}
\newcommand{\Hom}{\operatorname{Hom}}
\newcommand{\im}{\operatorname{im}}
\newcommand{\codim}{\operatorname{codim}}
\newcommand{\into}{\hookrightarrow}
\newcommand{\onto}{\to\!\!\!\!\!\to}
\newcommand{\ses}[3]{0\rightarrow #1\rightarrow #2\rightarrow#3\rightarrow 0}
\newcommand{\Sc}[2]{\langle #1,#2\rangle}
\newcommand{\spn}{\operatorname{span}}
\setcounter{MaxMatrixCols}{20}
\title{Cell Decompositions for Rank Two Quiver Grassmannians}

@@

\section{Introduction}
A cluster algebra is a subalgebra of a rational function field where generators are defined by an explicit recursion \cite{fomin-zelevinsky}.
Although the generators of the cluster algebra are given as rational functions they actually turn out to be Laurent polynomials.
These Laurent polynomial expressions for the cluster variables have been studied extensively and ultimately can be described using the representation theory of quivers \cite{caldero-chapoton,caldero-keller}.

-something about categorification and quiver Grassmannians
-something about compatible pairs and combinatorial construction of cluster variables
-statement of our results
-something about torus actions and the universal cover
-something about cell decompositions of quiver Grassmannians
-acknowledgements?

@@

\noindent Denote by $K(n)$ the \emph{$n$-Kronecker quiver} $1\stackrel{n}{\longleftarrow}2$ with vertices $K_0(n)=\{1,2\}$ and $n$ arrows from vertex $2$ to vertex $1$. 
The category $\rep K(n)$ of finite-dimensional representations of $K(n)$ is equivalent to the category of modules over the path algebra $A(n)$ of $K(n)$.
As a $\CC$-vector space, the path algebra $A(n)$ can be written as $A_0\oplus A_1$, where 
\begin{itemize}
  \item $A_0=\CC e_1\oplus \CC e_2$ is a two-dimensional semisimple algebra with orthogonal idempotents $e_1$ and $e_2$;
  \item $A_1=\bigoplus_{i=1}^n \CC a_i$ is the $A_0$-bimodule spanned by the arrows of $K(n)$, in particular $e_ka_ie_\ell=\delta_{k1}\delta_{\ell2}a_i$ for $1\le i\le n$ and $k,\ell\in\{1,2\}$.
\end{itemize}
Observe that $GL_n(\CC)$ naturally acts on $A_1$ and hence $GL_n(\CC)$ acts on $\rep K(n)$ via the induced action on $A(n)$.
More precisely, given a representation $M=(M_1,M_2,M_{a_i})$ of $K(n)$ and $g=(g_{ij})\in GL_n(\CC)$, the representation $g.M$ is given by $(M_1,M_2,M'_{a_i})$ with $M'_{a_i}=\sum\limits_{j=1}^n g_{ij}M_{a_j}$.
Note that $M$ and $g.M$ are not necessarily isomorphic as representations of $K(n)$.
\begin{lemma}
  For any morphism $\theta:M\to N$ between representations $M,N\in\rep K(n)$, the same maps $\theta_1:M_1\to N_1$ and $\theta_2:M_2\to N_2$ give a morphism $\theta:g.M\to g.N$ for any $g\in GL_n(\CC)$.
\end{lemma}
\begin{proof}
  Suppose $\theta:M\to N$ is a morphism of representations, i.e.\ $\theta_2\circ M_{a_j}=N_{a_j}\circ\theta_1$ for $1\le j\le n$.
  Then for $g=(g_{ij})\in GL_n(\CC)$, we have
  \[\theta_2\circ M'_{a_i}=\sum\limits_{j=1}^n g_{ij} \theta_2\circ M_{a_j}=\sum\limits_{j=1}^n g_{ij} M_{a_j}\circ\theta_1=N'_{a_i}\circ\theta_1\]
  for $1\le i\le n$ so that $\theta$ gives a morphism from $g.M$ to $g.N$. 
\end{proof}

Write $\Sigma_1$ and $\Sigma_2$ for the BGP-reflection functors of $K(n)$ \cite{bgp}. 
Note that (up to relabeling vertices) the quiver $K(n)$ itself does not change when applying the reflection functors and thus we view these as endofunctors acting on the category $\rep K(n)$ of finite-dimensional representations. %, but the vertices $1$ and $2$ are exchanged. 
With this identification of $K(n)$ and $K(n)^{op}$, the endofunctors $\Sigma_2\Sigma_1$ and $\Sigma_1\Sigma_2$ are each naturally isomorphic to the identity maps on the full subcategories $\rep_{\langle 1\rangle} K(n),\rep_{\langle 2\rangle} K(n)\subset \rep K(n)$, where the objects of $\rep_{\langle i\rangle} K(n)$ are those representations of $K(n)$ which do not contain the simple $S_i$ as a direct summand.
Also, following \cite{brenner-butler}, the Auslander-Reiten translation $\tau:\rep K(n)\to\rep K(n)$ may be identified with the functor $\Sigma_1^2$.

\begin{lemma}
  The reflection functors $\Sigma_i:\rep K(n)\to\rep K(n)$ are $GL_n(\CC)$-equivariant. 
\end{lemma}

Define Chebyshev polynomials $u_k$ for $k\in\ZZ$ by the recursion $u_0=0$, $u_1=1$, $u_{k+1}=nu_k-u_{k-1}$.
\begin{theorem}
  For each $m\ge1$, there exist unique (up to isomorphism) indecomposable rigid representations $P_m$ and $I_m$ of $K(n)$ with dimension vectors $(u_m,u_{m-1})$ and $(u_{m-1},u_m)$ respectively. 
  Moreover, any rigid representation of $K(n)$ is isomorphic to one of the form $P_m^{a_1}\oplus P_{m+1}^{a_2}$ or $I_m^{a_1}\oplus I_{m+1}^{a_2}$ for some $m\ge1$ and some $a_1,a_2\ge0$.
\end{theorem}
The representations $P_m$ are called the \emph{preprojective} representations of $K(n)$ and the representations $I_m$ are called \emph{preinjective}.
\begin{remark}
  \label{rem:reflection recursion}
  These representations satisfy the following recursions using the reflection functors:
  \[P_1=S_1,\quad P_m=\Sigma_2 P_{m-1},\quad I_1=S_2,\quad I_m=\Sigma_1 I_{m-1}\]
  for $m\ge2$.
  In particular, we have $P_{m-1}=\tau P_{m+1}$ and $I_{m+1}=\tau I_{m-1}$ for $m\ge2$. 
\end{remark}

Write $H_m:=\Hom(P_m,P_{m+1})$ for $m\ge1$.  
For each $m\ge2$, there is an Auslander-Reiten sequence \cite{auslander-reiten-smalo}
\begin{equation}
  \label{eq:AR sequence}
  0\longrightarrow P_{m-1}\stackrel{\iota_{m-1}}{\longrightarrow} P_m\otimes H_m\stackrel{ev}{\longrightarrow} P_{m+1}\longrightarrow 0,
\end{equation}
where the right-hand morphism is the natural evaluation map.

Write $Gr(H_m)$ for the \emph{total Grassmannian} of $H_m$ whose elements are nontrivial proper subspaces $V\subset H_m$.
Some results below remain true if we allow $H_m$ or $0$ as elements of $Gr(H_m)$, but not all, so for uniformity of exposition we omit these possibilities.
\begin{lemma}
  \label{le:injective evaluation maps}
  For any $V\in Gr(H_m)$, the natural evaluation map $ev_V:P_m\otimes V\to P_{m+1}$ is injective.
\end{lemma}

%%%%%%%%%%%%%%%%%%%
% needs polishing %
%%%%%%%%%%%%%%%%%%%
%\begin{proof}
%  A nonzero element $f\in V$ corresponds to an irreducible morphism $f:P_m\to P_{m+1}$ and thus does not factor through any representation $Z\ne P_m$, in particular $f$ is injective.
%  But any nonzero element of $\ker ev_V$ gives rise to a nonzero $f\in V$ which is not injective and thus $ev_V$ must be an injective map.
%\end{proof}

%\begin{lemma}
%  For any $m\ge1$, we have 
%  \[\dim\Hom(P_m,P_m)=1,\quad \dim\Hom(P_m,P_{m+1})=n,\quad \dim\Hom(P_{m+1},P_m)=0, \quad\dim\Ext(P_{m+1},P_m)=0.\]  
%  Moreover, for $1\le k<n$ and linearly independent functions $f_1,\ldots,f_k\in\Hom(P_m,P_{m+1})$, the map $[f_1\ \cdots\ f_k]:P_m^k\to P_{m+1}$ is injective. 
%\end{lemma}
%\begin{proof} 
%  Note that we have an Auslander-Reiten sequence
%  \begin{equation}
%    \label{eq:AR sequence}
%    0\longrightarrow P_{m-1}\stackrel{\iota_{m-1}}{\longrightarrow} P_m^n\stackrel{[f_1\ \cdots\ f_n]}{\longrightarrow} P_{m+1}\longrightarrow 0
%  \end{equation}
%  with irreducible homomorphisms $f_i$ which are also linearly independent. 
%  Thus they do not factor through a representation $Z\neq P_{m}$ which already means that they are injective, see for instance \cite[Lemma 1.6]{ass}. 
%  If we pick $k<n$ linearly independent homomorphisms $f_{i_1},\ldots, f_{i_k}$, this also means that $[f_{i_1}\ \cdots\ f_{i_k}]:P_m^k\to P_{m+1}$ is injective. 
%  Indeed otherwise, at least one of the homomorphisms would be forced to factor through a representation $Z\neq P_{m}$.
%\end{proof}
%\begin{remark}
%  Fix a basis $\{f_{m,1},\ldots,f_{m,n}\}$ for $\Hom(P_m,P_{m+1})$, $m\ge2$.
%  Define a basis $\{f_{\ell,1},\ldots,f_{\ell,n}\}$ for each $\Hom(P_\ell,P_{\ell+1})$, $1\le\ell\le m-1$, by downward induction as follows.  
%  Write $\iota_\ell:P_\ell\to P_{\ell+1}^n$ for the left-hand morphism from the sequence \eqref{eq:AR sequence} computed using $f_k=f_{\ell+1,k}$, $1\le k\le n$. 
%  Then the basis element $f_{\ell,k}$ is given by composing $\iota_\ell$ with the coordinate projection $\pi_k:P_{\ell+1}^n\to P_{\ell+1}$ onto the $k$-th factor.
%  Indeed, applying $\Hom(-,P_{\ell+1})$ to this sequence \eqref{eq:AR sequence} gives an isomorphism $\Hom(P_{\ell+1}^n,P_{\ell+1})\cong\Hom(P_\ell,P_{\ell+1})$ whereby the natural basis of coordinate projections on the left is taken to the basis $\{f_{\ell,k}\}$ on the right.
%\end{remark}
\begin{corollary}
  \label{cor:preprojective homomorphism duality}
  The sequence \eqref{eq:AR sequence} gives rise to an isomorphism $H_m^*\cong H_{m-1}$.
\end{corollary}
\begin{proof}
  Define a map $H_m^*\to H_{m-1}$ by $\varphi\mapsto\bar{\varphi}:=(id\overline{\otimes}\varphi)\circ\iota_{m-1}$, in words $\bar{\varphi}$ acts on $x\in P_{m-1}$ by contracting with the second factor in $\iota_{m-1}(x)$ to give an element of $P_m$.
  Suppose $\varphi\in H_m^*$ is a nonzero functional on $H_m$ and let $V\subsetneq H_m$ denote the kernel of $\varphi$.
  Then $\bar{\varphi}=0\in H_{m-1}$ if and only if the image of $\iota_{m-1}$ is contained in $P_m\otimes V\subsetneq P_m\otimes H_m$.
  But then Lemma~\ref{le:injective evaluation maps} implies $\iota_{m-1}$ is the zero map, a contradiction.
  Thus the map $H_m^*\to H_{m-1}$, $\varphi\mapsto\bar{\varphi}$ must be injective and hence an isomorphism.
\begin{remark}
  The set $Gr(H_m)$ is naturally a poset and Corollary~\ref{cor:preprojective homomorphism duality} gives an identification of the opposite poset $Gr(H_m)^{op}\cong Gr(H_m^*)$ with $Gr(H_{m-1})$.
  We write $\bar{V}\subset H_{m-1}$ for the subspace corresponding to $V\subset H_m$ under this identification.
  Under the isomorphism of $H_{m-1}$ with $H_m^*$, we have $\bar{V}=(H_m/V)^*$.
Following Remark~\ref{rem:reflection recursion}, the reflection functor $\Sigma_1$ provides a natural identification of $H_m$ with $H_{m-1}$ for $m\ge2$. 
Combining with Corollary~\ref{cor:preprojective homomorphism duality} we obtain a natural isomorphism $H_m^*\cong H_m$ for $m\ge2$.
The next observation is essential for the results to follow.
\begin{proposition}
  For $V\in Gr(H_m)$, $m\ge1$, we have $\bar{\bar{V}}=\tau V$.
\end{proposition}
\begin{proof}
  We show that $\bar{\bar{\varphi}}=\tau\varphi$ for $\varphi\in H_m$.
  Indeed, we have
  \[\bar{\bar{\varphi}}=(id\overline{\otimes}\bar{\varphi})\circ\iota_{m-2}=\Big(id\overline{\otimes}\big((id\overline{\otimes}\varphi)\circ\iota_{m-1}\big)\Big)\circ\iota_{m-2}.\]

\begin{definition}
  For $V\in Gr(H_m)$, define the \emph{truncated preprojective} $P_{m+1}^V$ to be the cokernel of the map $ev_V:P_m\otimes V\to P_{m+1}$, i.e.\ we have a short exact sequence
  \begin{equation}
    \label{eq:truncated preprojectives}
    0\longrightarrow P_m\otimes V\stackrel{ev_V}{\longrightarrow} P_{m+1}\stackrel{\pi_V}{\longrightarrow} P_{m+1}^V\longrightarrow 0.
  \end{equation}
\end{definition}
\begin{remark}
  It will be convenient to also set $P_{m+1}^0=P_{m+1}$, observe that this notation is consistent with taking $V=0$ in the sequence \eqref{eq:truncated preprojectives}.
\end{remark}

>>>>

\newcommand{\bfe}{\mathbf{e}}
\newcommand{\bff}{\mathbf{f}}
\newcommand{\bfg}{\mathbf{g}}
\newcommand{\bfi}{\mathbf{i}}
\newcommand{\bfh}{\mathbf{h}}
\newcommand{\tbfe}{{\tilde\bfe}}
\newcommand{\tbff}{{\tilde\bff}}
\newcommand{\tbfg}{{\tilde\bfg}}
\newcommand{\tbfh}{{\tilde\bfh}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\cC}{\mathcal{C}}
\newcommand{\cG}{\mathcal{G}}
\newcommand{\cP}{\mathcal{P}}
\newcommand{\ui}{\underline i}
\newcommand{\uj}{\underline j}
\newcommand\udim{{\underline{\dim}\, }}
\newcommand{\pt}{\mathrm{pt}}
\newcommand{\Rep}{\mathrm{Rep}}
\newcommand{\rep}{\operatorname{rep}}
\newcommand{\Gr}{\mathrm{Gr}}
\renewcommand{\AA}{\mathbb{A}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\kk}{\Bbbk}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\Ext}{\operatorname{Ext}}
\newcommand{\Hom}{\operatorname{Hom}}
\newcommand{\im}{\operatorname{im}}
\newcommand{\codim}{\operatorname{codim}}
\newcommand{\into}{\hookrightarrow}
\newcommand{\onto}{\to\!\!\!\!\!\to}
\newcommand{\ses}[3]{0\rightarrow #1\rightarrow #2\rightarrow#3\rightarrow 0}
\newcommand{\Sc}[2]{\langle #1,#2\rangle}
\newcommand{\spn}{\operatorname{span}}
\setcounter{MaxMatrixCols}{20}

\hyphenation{endo-functors}

\title{Cell Decompositions for Rank Two Quiver Grassmannians}

@@

\section{Introduction}
A cluster algebra is a subalgebra of a rational function field where generators are defined by an explicit recursion \cite{fomin-zelevinsky}.
Although the generators of the cluster algebra are given as rational functions they actually turn out to be Laurent polynomials.
These Laurent polynomial expressions for the cluster variables have been studied extensively and ultimately can be described using the representation theory of quivers \cite{caldero-chapoton,caldero-keller}: the cluster variables are generating functions for Euler characteristics of quiver Grassmannians.

For rank two cluster algebras, the Laurent expressions of cluster variables can also be computed using a certain Dyck path combinatorics \cite{lee-li-zelevinsky}.
The confluence of these results gives rise to a combinatorial construction for the Euler characteristics of certain quiver Grassmannians.
Our main result is a geometric explanation for this correspondence by constructing a decomposition of the quiver Grassmannians into affine cells which are labeled by compatible pairs.

-something about compatible pairs and combinatorial construction of cluster variables
-statement of our results
-something about torus actions and the universal cover
-something about cell decompositions of quiver Grassmannians
-acknowledgements?

@@

\noindent Denote by $K(n)$ the \emph{$n$-Kronecker quiver} $1\stackrel{n}{\longleftarrow}2$ with vertices $K_0(n)=\{1,2\}$ and $n$ arrows from vertex $2$ to vertex $1$. 
The category $\rep K(n)$ of finite-dimensional representations of $K(n)$ is equivalent to the category of modules over the path algebra $A(n)$ of $K(n)$.
As a $\CC$-vector space, the path algebra $A(n)$ can be written as $A_0\oplus A_1$, where 
\begin{itemize}
  \item $A_0=\CC e_1\oplus \CC e_2$ is a two-dimensional semisimple algebra with orthogonal idempotents $e_1$ and $e_2$;
  \item $A_1=\bigoplus_{i=1}^n \CC a_i$ is the $A_0$-bimodule spanned by the arrows of $K(n)$, that is $e_ka_ie_\ell=\delta_{k1}\delta_{\ell2}a_i$ for $1\le i\le n$ and $k,\ell\in\{1,2\}$.
\end{itemize}
Observe that $GL_n(\CC)$ naturally acts on $A_1$ and hence $GL_n(\CC)$ acts on $\rep K(n)$ via the induced action on $A(n)$.
More precisely, given a representation $M=(M_1,M_2,M_{a_i})$ of $K(n)$ and $g=(g_{ij})\in GL_n(\CC)$, the representation $g.M$ is given by $(M_1,M_2,M'_{a_i})$ with $M'_{a_i}=\sum\limits_{j=1}^n g_{ij}M_{a_j}$.
Note that $M$ and $g.M$ are not necessarily isomorphic as representations of $K(n)$.
\begin{lemma}
  \label{le:hom equivariance}
  For any morphism $\theta:M\to N$ between representations $M,N\in\rep K(n)$, the same maps $\theta_1:M_1\to N_1$ and $\theta_2:M_2\to N_2$ give a morphism $\theta^g:g.M\to g.N$ for any $g\in GL_n(\CC)$.
  In particular, the hom-spaces $\Hom(M,N)$ and $\Hom(g.M,g.N)$ are canonically identified for each $g\in G$.
\end{lemma}
\begin{proof}
  Suppose $\theta:M\to N$ is a morphism of representations, i.e.\ $\theta_2\circ M_{a_j}=N_{a_j}\circ\theta_1$ for $1\le j\le n$.
  Then for $g=(g_{ij})\in GL_n(\CC)$, we have
  \[\theta_2\circ M'_{a_i}=\sum\limits_{j=1}^n g_{ij}(\theta_2\circ M_{a_j})=\sum\limits_{j=1}^n g_{ij}(M_{a_j}\circ\theta_1)=N'_{a_i}\circ\theta_1\]
  for $1\le i\le n$ so that $\theta$ also gives a morphism from $g.M$ to $g.N$. 
\end{proof}
\begin{corollary}
  \label{cor:indecomposability}
  For $M\in\rep K(n)$ and $g\in GL_n(\CC)$, the representation $M$ is indecomposable if and only if $g.M$ is indecomposable.
\end{corollary}
\begin{proof}
  The representation $M$ is decomposable if there exists a split epimorphism $\theta:M\onto N$ for some nonzero representation $N$.
  But this occurs exactly when the map $\theta^g:g.M\onto g.N$ is also a split epimorphism.
\end{proof}

Write $\Sigma_1$ and $\Sigma_2$ for the BGP-reflection functors of $K(n)$ \cite{bgp}. 
Note that (up to relabeling vertices) the quiver $K(n)$ itself does not change when applying the reflection functors and thus we view these as endofunctors acting on the category $\rep K(n)$. 
With this identification of $K(n)$ and $K(n)^{op}$, the endofunctors $\Sigma_2\Sigma_1$ and $\Sigma_1\Sigma_2$ are each naturally isomorphic to the identity maps on the full subcategories $\rep_{\langle 1\rangle} K(n),\rep_{\langle 2\rangle} K(n)\subset \rep K(n)$, where the objects of $\rep_{\langle i\rangle} K(n)$ are those representations of $K(n)$ which do not contain the simple $S_i$ as a direct summand.
Also, following \cite{brenner-butler}, the Auslander-Reiten translation $\tau:\rep K(n)\to\rep K(n)$ may be identified with the functor $\Sigma_1^2$.

\begin{lemma}
  \footnote{Is this true?}
  The reflection functors $\Sigma_i:\rep K(n)\to\rep K(n)$ are $GL_n(\CC)$-equivariant. 
\end{lemma}

Define Chebyshev polynomials $u_k$ for $k\in\ZZ$ by the recursion $u_0=0$, $u_1=1$, $u_{k+1}=nu_k-u_{k-1}$.
The following is well-known.
\begin{theorem}
  \label{th:rigids}
  For each $m\ge1$, there exist unique (up to isomorphism) indecomposable rigid representations $P_m$ and $I_m$ of $K(n)$ with dimension vectors $(u_m,u_{m-1})$ and $(u_{m-1},u_m)$ respectively. 
  Moreover, any rigid representation of $K(n)$ is isomorphic to one of the form $P_m^{a_1}\oplus P_{m+1}^{a_2}$ or $I_m^{a_1}\oplus I_{m+1}^{a_2}$ for some $m\ge1$ and some $a_1,a_2\ge0$.
\end{theorem}
The representations $P_m$ are called the \emph{preprojective} representations of $K(n)$ and the representations $I_m$ are called \emph{preinjective}.
\begin{remark}
  \label{rem:reflection recursion}
  These representations satisfy the following recursions using the reflection functors:
  \[P_1=S_1,\quad P_m=\Sigma_2 P_{m-1},\quad I_1=S_2,\quad I_m=\Sigma_1 I_{m-1}\]
  for $m\ge2$.
  In particular, we have $P_{m-1}=\tau P_{m+1}$ and $I_{m+1}=\tau I_{m-1}$ for $m\ge2$. 
\end{remark}
\begin{corollary}
  \label{cor:preprojective stability}
  For $m\ge1$ and $g\in G$, the representations $P_m$ and $g.P_m$ are isomorphic.
\end{corollary}
\begin{proof}
  By construction, the representation $g.P_m$ also has dimension vector $(u_m,u_{m-1})$ and is indecomposable by Corollary~\ref{cor:indecomposability}.
  Thus by Theorem~\ref{th:rigids}, $P_m$ and $g.P_m$ are isomorphic.
\end{proof}
For $g\in GL_n(\CC)$, write $\phi_{m,g}:P_m\to g.P_m$ for the (unique up to scalars) isomorphism from Corollary~\ref{cor:preprojective stability}.

Set $H_m:=\Hom(P_m,P_{m+1})$ for $m\ge1$ and define an action of $GL_n(\CC)$ on $H_m$ by $g.\theta:=\phi_{m+1,g}^{-1}\circ\theta^g\circ\phi_{m,g}$.
Note that the first $\theta$ here represents an element of $H_m$ while the second $\theta^g$ represents the corresponding element of $H_m^g:=\Hom(g.P_m,g.P_{m+1})$ from Lemma~\ref{le:hom equivariance}.

Write $Gr(H_m)$ for the \emph{total Grassmannian} of $H_m$ whose elements are nontrivial proper subspaces $V\subset H_m$.
Some results below remain true if we allow $H_m$ or $0$ as elements of $Gr(H_m)$, but not all, so for uniformity of exposition we omit these possibilities.
Lemma~\ref{le:hom equivariance} gives an identification of $Gr(H_m)$ with $Gr(H_m^g)$ and we write $V\mapsto V^g$ for this correspondence.
For $1\le d\le n-1$, let $Gr_d(H_m)\subset Gr(H_m)$ be the subvariety of $d$-dimensional subspaces of $H_m$.
\begin{lemma}
  \label{le:transitivity}
  The action of $GL_n(\CC)$ on $H_m$ induces a transitive action of $GL_n(\CC)$ on each component $Gr_d(H_m)$.
\end{lemma} 
\begin{proof}

\end{proof}
\begin{lemma}
  \label{le:injective evaluation maps}
  For any $V\in Gr(H_m)$, the natural evaluation map $ev_V:P_m\otimes V\to P_{m+1}$ is injective.
\end{lemma}
%%%%%%%%%%%%%%%%%%%
% needs polishing %
%%%%%%%%%%%%%%%%%%%
\begin{proof}
  \footnote{This needs to be cleaned up, the proof seems false as written.}
  A nonzero element $f\in V$ corresponds to an irreducible morphism $f:P_m\to P_{m+1}$ and thus does not factor through any representation $Z\ne P_m$, in particular $f$ is injective.
  But any nonzero element of $\ker ev_V$ gives rise to a nonzero $f\in V$ which is not injective and thus $ev_V$ must be an injective map.
\end{proof}

In what follows we will not distinguish between $P_m\otimes V$ and its image under $ev_V$.
Similarly, for $V^g\in Gr(H_m^g)$, we identify $g.P_m\otimes V^g$ with its image in $g.P_{m+1}$ under the natural evaluation map. 
\begin{lemma}
  \label{le:g equalities}
  For any $g\in GL_n(\CC)$, the subrepresentation $g.(P_m\otimes V)\subset g.P_{m+1}$ is equal to $g.P_m\otimes V^g\subset g.P_{m+1}$ and the subrepresentation $\phi_{m+1,g}^{-1}(g.P_m\otimes V^g)\subset P_{m+1}$ is equal to $P_m\otimes g.V\subset P_{m+1}$.
\end{lemma}
\begin{proof}
  The first equality is immediate from the definitions.
  The second equality also follows from the definitions by observing that $\phi_{m+1,g}(P_m\otimes g.V)=g.P_m\otimes V^g$.
\end{proof}

Following Lemma~\ref{le:injective evaluation maps} we define the following.
\begin{definition}
  For $V\in Gr(H_m)$, define the \emph{truncated preprojective} $P_{m+1}^V$ to be the cokernel of the map $ev_V:P_m\otimes V\to P_{m+1}$, i.e.\ we have a short exact sequence
  \begin{equation}
    \label{eq:truncated preprojectives}
    0\longrightarrow P_m\otimes V\stackrel{ev_V}{\longrightarrow} P_{m+1}\stackrel{\pi_V}{\longrightarrow} P_{m+1}^V\longrightarrow 0.
  \end{equation}
\end{definition}
\begin{remark}
  It will be convenient to also set $P_{m+1}^0=P_{m+1}$, observe that this notation is consistent with taking $V=0$ in the sequence \eqref{eq:truncated preprojectives}.
\end{remark}

Any subrepresentation $E\subset P_{m+1}^V$ corresponds by taking pullbacks to a chain of inclusions $P_m\otimes V\subset\tilde E\subset P_{m+1}$.
\begin{theorem}
  The quiver Grassmannian $Gr_\bfe(P_{m+1}^V)$ is smooth for each dimension vector $\bfe$ and each $V\in Gr(H_m)$.
  Moreover, for $V,W\in Gr_d(H_m)$, we have $Gr_\bfe(P_{m+1}^V)\cong Gr_\bfe(P_{m+1}^W)$ for all dimension vectors $\bfe$.
\end{theorem}
\begin{proof}
  Both claims are proven via the same geometric construction.
  For each dimension vector $\bfe$, there is a smooth variety
  \[X_{\bfe,m}=\{(\tilde E,V):V\in Gr_d(H_m),\ P_m\otimes V\subset\tilde E\subset P_{m+1}i,\ \udim\tilde E=\bfe+d\udim P_m\}.\]
  Under the natural projection $\pi:X_{\bfe,m}\to Gr_d(H_m)$, the fiber $\pi^{-1}(V)$ identifies with $Gr_\bfe(P_{m+1}^V)$.
  In particular, this implies $Gr_\bfe(P_{m+1}^V)$ is smooth for each $V\in Gr_d(H_m)$.

  The variety $X_{\bfe,m}$ supports an action of $GL_n(\CC)$ by setting $g.(\tilde E,V)=\big(\phi_{m+1,g}^{-1}(g.\tilde E),g.V\big)$.
  Indeed, by Lemma~\ref{le:g equalities} we have 
  \[P_m\otimes g.V=\phi_{m+1,g}^{-1}(g.P_m\otimes V^g)=\phi_{m+1,g}^{-1}\big(g.(P_m\otimes V)\big)\subset\phi_{m+1,g}^{-1}(g.\tilde E).\]
  Observe that the projection map $\pi$ is $GL_n(\CC)$-equivariant and thus by Lemma~\ref{le:transitivity} the quiver Grassmannians $Gr_\bfe(P_{m+1}^V)$ and $Gr_\bfe(P_{m+1}^W)$ are isomorphic for any $V,W\in Gr_d(H_m)$.
\end{proof}

%\begin{lemma}
%  For any $m\ge1$, we have 
%  \[\dim\Hom(P_m,P_m)=1,\quad \dim\Hom(P_m,P_{m+1})=n,\quad \dim\Hom(P_{m+1},P_m)=0, \quad\dim\Ext(P_{m+1},P_m)=0.\]  
%  Moreover, for $1\le k<n$ and linearly independent functions $f_1,\ldots,f_k\in\Hom(P_m,P_{m+1})$, the map $[f_1\ \cdots\ f_k]:P_m^k\to P_{m+1}$ is injective. 
%\end{lemma}
%\begin{proof} 
%  Note that we have an Auslander-Reiten sequence
%  \begin{equation}
%    \label{eq:AR sequence}
%    0\longrightarrow P_{m-1}\stackrel{\iota_{m-1}}{\longrightarrow} P_m^n\stackrel{[f_1\ \cdots\ f_n]}{\longrightarrow} P_{m+1}\longrightarrow 0
%  \end{equation}
%  with irreducible homomorphisms $f_i$ which are also linearly independent. 
%  Thus they do not factor through a representation $Z\neq P_{m}$ which already means that they are injective, see for instance \cite[Lemma 1.6]{ass}. 
%  If we pick $k<n$ linearly independent homomorphisms $f_{i_1},\ldots, f_{i_k}$, this also means that $[f_{i_1}\ \cdots\ f_{i_k}]:P_m^k\to P_{m+1}$ is injective. 
%  Indeed otherwise, at least one of the homomorphisms would be forced to factor through a representation $Z\neq P_{m}$.
%\end{proof}
%\begin{remark}
%  Fix a basis $\{f_{m,1},\ldots,f_{m,n}\}$ for $\Hom(P_m,P_{m+1})$, $m\ge2$.
%  Define a basis $\{f_{\ell,1},\ldots,f_{\ell,n}\}$ for each $\Hom(P_\ell,P_{\ell+1})$, $1\le\ell\le m-1$, by downward induction as follows.  
%  Write $\iota_\ell:P_\ell\to P_{\ell+1}^n$ for the left-hand morphism from the sequence \eqref{eq:AR sequence} computed using $f_k=f_{\ell+1,k}$, $1\le k\le n$. 
%  Then the basis element $f_{\ell,k}$ is given by composing $\iota_\ell$ with the coordinate projection $\pi_k:P_{\ell+1}^n\to P_{\ell+1}$ onto the $k$-th factor.
%  Indeed, applying $\Hom(-,P_{\ell+1})$ to this sequence \eqref{eq:AR sequence} gives an isomorphism $\Hom(P_{\ell+1}^n,P_{\ell+1})\cong\Hom(P_\ell,P_{\ell+1})$ whereby the natural basis of coordinate projections on the left is taken to the basis $\{f_{\ell,k}\}$ on the right.
%\end{remark}
